### 2024-05-04 paper/sec/3_methods.tex
3.3 detecting poses and gestures

Thus, we decided to add some math equations to transform our features from absolute x, y positions to angles between detected 
key points. Additionally, this will likely aid the model in recognizing elliptical motions, which are prevalent in physical 
therapy exercises, as a continuous motion, as opposed to a given, singular elliptical motion being calculated as multiple 
straight-line movements. Additionally, this encodes information about body parts relative to each

>>>>

Thus, we decided to add some math equations to transform our features from absolute x, y positions to angles between detected 
key points. Additionally, this will likely aid the model in recognizing elliptical motions, which are prevalent in physical 
therapy exercises, as a continuous motion, as opposed to a given, singular elliptical motion being calculated as multiple 
straight-line movements. Additionally, this encodes information about body parts relative to each other. For 
instance, moving one’s right hand downwards on the left side of one’s body could be differentiated from moving 
one’s right hand downwards on the right side of one’s body through the angle of the right elbow relative to 
the right shoulder.

\subsection{Pose and Gesture Extraction}
By the temporal nature of videos and humans’ inability to teleport, many frames have body angles that are 
close in value to those of the neighboring frames, similar to how in images, many pixels may have color 
values close to their neighbors. It would be inefficient to always work with all of the frames in a given 
video. In fact, in images, features are often only extracted for corners. We suggest the idea of extracting 
poses and gestures from videos.

While the algorithm we’ve designed is inspired by the Harris Corner Detector, poses and gestures differ from 
corners in that we’re not looking for a gradient across all dimensions. Intuitively, this makes sense: even 
if a person were to stand-in place (not change the angles in their lower body), we would still consider a 
hand-wave to be a gesture, and raising different hands would be considered different poses.

To reiterate, unlike a corner, which only has two dimensions x and y, a pose or gesture would have as many 
dimensions as there are angles. Additionally, while corners are detected based on the ratio of the 
eigenvalues, we’re more concerned with having one strong edge or many weak edges in identifying a gesture: 
requiring all angles to change would omit basic gestures like a simple hand wave because the majority of 
angles in the body wouldn’t change. Hence, we use Euclidean Distance as a “change in body position” score.

Another point of interest is that poses would have a low “change in body position” score. Going back to the 
corner detection analogy, poses are analogous to a flat area within an  image, yet they arguably define a 
person’s physical actions more than the brief moments of highest “change in body position”. Thus, to extract 
poses from angles, we actually apply non-minimum suppression (instead of the Harris Corner Detector’s 
non-maximum suppression) after calculating the Euclidean Distance between angles in adjacent frames.

Meanwhile, we consider a gesture to be the opposite of a pose, a small range of frames when there is a 
significant “change in body position”. Arguably, a gesture would take place for all the frames between 
two poses, not just for moments of the fastest “change in body position”. However, because we applied a 
threshold that selects only a subset of the most significant poses, the above definition may cause multiple 
gestures to be recognized as one, so we have currently defined a gesture to be three consecutive frames of 
significant change in angles. (However, it could be interesting to explore modified algorithms for detecting 
poses and gestures in future work.)
